---
title: Monte Carlo Simulation for City of Flagstaff Home Weatherization and Energy Rebate Program
author: Brandon McNamara
project: Dissertation chapter 3
date: March 11, 2025
---

Package Install & Load
```{r setup}
if (!require("dplyr")) install.packages("dplyr")
library(dplyr)
if (!require("readxl")) install.packages("readxl")
library(readxl)
if (!require("writexl")) install.packages("writexl")
library(writexl)
if (!require("ggplot2")) install.packages("ggplot2")
library(ggplot2)
if (!require("scales")) install.packages("scales")
library(scales)
if (!require("data.table")) install.packages("data.table")
library(data.table)
if (!require("gridExtra")) install.packages("gridExtra")
library(gridExtra)
```

Set up global variables
```{r}
# Define the following global variables to be called by parameter functions
run_count <- 1000
scenario_start <- 2026
intv_years <- 10 # The number of years the intervention will be implemented
intv_lifetime <- 15 # The number of years the intervention generates benefits, generally the expected lifetime of equipment
scenario_end  <- scenario_start + intv_years + intv_lifetime - 1
scenario_years <- scenario_end - scenario_start + 1
output_headers <- scenario_start:scenario_end
```

Initiate helper functions
```{r}
source("parameter_functions.r")
```

Create CO2e Emission Factor Library
```{r}
EFL_co2e("ar6")
write_xlsx(EFL, path = "EFLco2e.xlsx") 
```

Emission Factor Simulations
```{r}
# Electricity emission factor
scenario_MCS("ef_sim_electricity", sheet_name = "electricity_ef", units = "kgco2e/kWh") #write a custome function for electricity? with the starting value as a input? And NREL grid region as an input as well. 

# Supply Chain emission factor
continuous_uniform2("ef_sim_appliances_upstream", minvalue = 0.079, maxvalue = 0.161, units = "kgco2e/USD")
continuous_uniform2("ef_sim_weatherize_upstream", minvalue = 0.082, maxvalue = 0.208, units = "kgco2e/USD")

```

Boundary Simulations
```{r}
# Reference boundary represents zero energy efficiency upgrades in each year
static_value("ref_boundary_mgnl", 0, units = "project count")

# Intervention boundary is modeled as a random value between 120 and 160.
# The program is assumed to have funding for 10 years. After that, no more homes are funded. 
discrete_uniform3("intv_boundary_mgnl", minvalue = 120, maxvalue = 160, nyears = intv_years, value2 = 0, units = "project count")

# Project types are randomly assigned for the intervention boundary, with three new dataframes for each project type.
assign_projecttype(intv_boundary_mgnl)

```

Air source heat pumps (ASHP): Activity data, cost, and GHG simulations
```{r}
# Simulate annual electricity consumption for space heating and cooling. Each home in the boundary is simulated individually. The intervention factor is the % energy savings, also simulated for each home in the boundary. 
calc_ad1("ashp_ref_ann_data1",
         "ashp_intv_ann_data1",
         ashp_boundary_mgnl, 
         ref_pdf = rnorm,
         ref_params = list(mean = 4161, sd = 233), # Reference electricity consumption is modeled with a normal distribution
         intv_pdf = runif, 
         intv_params = list(min = 0.31, max = 0.47), # Intervention factor, % savings, is modeled with a uniform distribution  
         units = "kWh")

# Calculate cumulative activity data
cumul_sum2("ashp_ref_data1", ashp_ref_ann_data1)
cumul_sum2("ashp_intv_data1", ashp_intv_ann_data1)

# Simulate annual natural gas consumption for space heating. Each home in the boundary is simulated individually. The intervention factor is the % energy savings, also simulated for each home in the boundary. 
calc_ad1("ashp_ref_ann_data2",
         "ashp_intv_ann_data2",
         ashp_boundary_mgnl, 
         ref_pdf = rnorm,
         ref_params = list(mean = 603, sd = 11.3), # Reference natural gas consumption is modeled with a normal distribution
         intv_pdf = runif, 
         intv_params = list(min = 0.31, max = 0.47), # Intervention factor, % savings,is modeled with a uniform distribution  
         units = "therms")

# Calculate cumulative activity data
cumul_sum2("ashp_ref_data2", ashp_ref_ann_data2)
cumul_sum2("ashp_intv_data2", ashp_intv_ann_data2)

# ASHP Capex is calculated by simulating the installation costs for each home
calculate_capex("ashp_ref_capex1", "ashp_intv_capex1", ref_boundary_mgnl, ashp_boundary_mgnl, capex_min = 2100, capex_max = 6000, units = "USD")

# Sum total Capex
ashp_ref_sumcapex <- ashp_ref_capex1[,-1]
ashp_intv_sumcapex <- ashp_intv_capex1[,-1]

# ASHP Opex is calculated for electricity and natural gas separately, based on current energy prices
calculate_opex("ashp_ref_opex1", "ashp_intv_opex1", ashp_ref_data1, ashp_intv_data1, opex = 0.1542, units = "USD")
calculate_opex("ashp_ref_opex2", "ashp_intv_opex2", ashp_ref_data2, ashp_intv_data2, opex = 0.3434, units = "therms")

# Sum total opex
ashp_ref_sumopex <- ashp_ref_opex1[,-1] + ashp_ref_opex2[,-1]
ashp_intv_sumopex <- ashp_intv_opex1[,-1] + ashp_intv_opex2[,-1]

# Calculate scope 1 natural gas emissions 
ghg_conversion1("ashp_ref_ghg_natgas",
                "ashp_intv_ghg_natgas",
                ashp_ref_data2,
                ashp_intv_data2,
                fltr.ef_activeyear = 2024, 
                fltr.service_type = "natural gas", 
                fltr.emission_category = "stationary", 
                fltr.emission_scope = "scope 1",
                fltr.service_subcategory1 = "", 
                fltr.service_subcategory2 = "", 
                fltr.country = "global", 
                fltr.subregion = "")

# Calculate scope 2 electricity ghg emissions based on a simulation of grid emissions intensity
ghg_conversion2("ashp_ref_ghg_elec", "ashp_intv_ghg_elec", ef_sim_electricity, ashp_ref_data1, ashp_intv_data1)

# Calculate scope 3 supply chain ghg emissions based on a simulation of the upstream spend-based emission factor
ghg_conversion3("ashp_ref_ghg_upstream", "ashp_intv_ghg_upstream", ef_sim_appliances_upstream, ashp_ref_capex1, ashp_intv_capex1)

# Calculate total ashp GHG emissions
ashp_ref_ghg_total <- ashp_ref_ghg_natgas[, -1] + ashp_ref_ghg_elec[, -1] + ashp_ref_ghg_upstream[, -1]
ashp_intv_ghg_total <- ashp_intv_ghg_natgas[, -1] + ashp_intv_ghg_elec[, -1] + ashp_intv_ghg_upstream[, -1]

```

Weatherization: Activity data, cost, and GHG simulations
```{r}
# Simulate annual electricity consumption for space heating and cooling. Each home in the boundary is simulated individually. The intervention factor is the % energy savings, also simulated for each home in the boundary. 
calc_ad1("weatherize_ref_ann_data1",
         "weatherize_intv_ann_data1",
         weatherize_boundary_mgnl, 
         ref_pdf = rnorm,
         ref_params = list(mean = 4161, sd = 233), # Reference electricity consumption is modeled with a normal distribution
         intv_pdf = runif, 
         intv_params = list(min = 0.10, max = 0.20), # Intervention factor, % savings, is modeled with a uniform distribution  
         units = "kWh")

# Calculate cumulative activity data
cumul_sum2("weatherize_ref_data1", weatherize_ref_ann_data1)
cumul_sum2("weatherize_intv_data1", weatherize_intv_ann_data1)

# Simulate annual natural gas consumption for space heating. Each home in the boundary is simulated individually. The intervention factor is the % energy savings, also simulated for each home in the boundary. 
calc_ad1("weatherize_ref_ann_data2",
         "weatherize_intv_ann_data2",
         weatherize_boundary_mgnl, 
         ref_pdf = rnorm,
         ref_params = list(mean = 603, sd = 11.3), # Reference natural gas consumption is modeled with a normal distribution
         intv_pdf = runif, 
         intv_params = list(min = 0.10, max = 0.20), # Intervention factor, % savings,is modeled with a uniform distribution  
         units = "therms")

# Calculate cumulative activity data
cumul_sum2("weatherize_ref_data2", weatherize_ref_ann_data2)
cumul_sum2("weatherize_intv_data2", weatherize_intv_ann_data2)

# Weatherization Capex is calculated by simulating the installation costs for each home
calculate_capex("weatherize_ref_capex1", "weatherize_intv_capex1", ref_boundary_mgnl, weatherize_boundary_mgnl, capex_min = 1879, capex_max = 2962, units = "USD")

# Sum total Capex
weatherize_ref_sumcapex <- weatherize_ref_capex1[,-1]
weatherize_intv_sumcapex <- weatherize_intv_capex1[,-1]

# Weatherization Opex is calculated for electricity and natural gas separately, based on current energy prices
calculate_opex("weatherize_ref_opex1", "weatherize_intv_opex1", weatherize_ref_data1, weatherize_intv_data1, opex = 0.1542, units = "USD")
calculate_opex("weatherize_ref_opex2", "weatherize_intv_opex2", weatherize_ref_data2, weatherize_intv_data2, opex = 0.3434, units = "therms")

# Sum total Opex
weatherize_ref_sumopex <- weatherize_ref_opex1[,-1] + weatherize_ref_opex2[,-1]
weatherize_intv_sumopex <- weatherize_intv_opex1[,-1] + weatherize_intv_opex2[,-1]

# Calculate scope 1 natural gas emissions 
ghg_conversion1("weatherize_ref_ghg_natgas",
                "weatherize_intv_ghg_natgas",
                weatherize_ref_data2,
                weatherize_intv_data2,
                fltr.ef_activeyear = 2024, 
                fltr.service_type = "natural gas", 
                fltr.emission_category = "stationary", 
                fltr.emission_scope = "scope 1",
                fltr.service_subcategory1 = "", 
                fltr.service_subcategory2 = "", 
                fltr.country = "global", 
                fltr.subregion = "")

# Calculate scope 2 electricity ghg emissions based on a simulation of grid emissions intensity
ghg_conversion2("weatherize_ref_ghg_elec", "weatherize_intv_ghg_elec", ef_sim_electricity, weatherize_ref_data1, weatherize_intv_data1)

# Calculate scope 3 supply chain ghg emissions based on a simulation of the upstream spend-based emission factor
ghg_conversion3("weatherize_ref_ghg_upstream", "weatherize_intv_ghg_upstream", ef_sim_weatherize_upstream, weatherize_ref_capex1, weatherize_intv_capex1)

# Calculate total weatherize GHG emissions
weatherize_ref_ghg_total <- weatherize_ref_ghg_natgas[, -1] + weatherize_ref_ghg_elec[, -1] + weatherize_ref_ghg_upstream[, -1]
weatherize_intv_ghg_total <- weatherize_intv_ghg_natgas[, -1] + weatherize_intv_ghg_elec[, -1] + weatherize_intv_ghg_upstream[, -1]

```

Heat Pump Water Heater (HPWH): Activity data, cost, and GHG simulations
```{r}
# Simulate annual electricity consumption for water heating. Each home in the boundary is simulated individually. The intervention factor is the % energy savings, also simulated for each home in the boundary. 
calc_ad1("hpwh_ref_ann_data1",
         "hpwh_intv_ann_data1",
         hpwh_boundary_mgnl, 
         ref_pdf = rnorm,
         ref_params = list(mean = 2969, sd = 108.1), # Reference electricity consumption is modeled with a normal distribution
         intv_pdf = runif, 
         intv_params = list(min = 0.30, max = 0.70), # Intervention factor, % savings, is modeled with a uniform distribution  
         units = "kWh")

# Calculate cumulative activity data
cumul_sum2("hpwh_ref_data1", hpwh_ref_ann_data1)
cumul_sum2("hpwh_intv_data1", hpwh_intv_ann_data1)

# Simulate annual natural gas consumption for space heating. Each home in the boundary is simulated individually. The intervention factor is the % energy savings, also simulated for each home in the boundary. 
calc_ad1("hpwh_ref_ann_data2",
         "hpwh_intv_ann_data2",
         hpwh_boundary_mgnl, 
         ref_pdf = rnorm,
         ref_params = list(mean = 237, sd = 4), # Reference natural gas consumption is modeled with a normal distribution
         intv_pdf = runif, 
         intv_params = list(min = 0.30, max = 0.70), # Intervention factor, % savings,is modeled with a uniform distribution  
         units = "therms")

# Calculate cumulative activity data
cumul_sum2("hpwh_ref_data2", hpwh_ref_ann_data2)
cumul_sum2("hpwh_intv_data2", hpwh_intv_ann_data2)

# HPWH Capex is calculated by simulating the installation costs for each home
calculate_capex("hpwh_ref_capex1", "hpwh_intv_capex1", ref_boundary_mgnl, hpwh_boundary_mgnl, capex_min = 1300, capex_max = 2900, units = "USD")

# Sum total Capex
hpwh_ref_sumcapex <- hpwh_ref_capex1[,-1]
hpwh_intv_sumcapex <- hpwh_intv_capex1[,-1]

# HPWH Opex is calculated for electricity and natural gas separately, based on current energy prices
calculate_opex("hpwh_ref_opex1", "hpwh_intv_opex1", hpwh_ref_data1, hpwh_intv_data1, opex = 0.1542, units = "USD")
calculate_opex("hpwh_ref_opex2", "hpwh_intv_opex2", hpwh_ref_data2, hpwh_intv_data2, opex = 0.3434, units = "therms")

# Sum total Opex
hpwh_ref_sumopex <- hpwh_ref_opex1[,-1] + hpwh_ref_opex2[,-1]
hpwh_intv_sumopex <- hpwh_intv_opex1[,-1] + hpwh_intv_opex2[,-1]

# Calculate scope 1 natural gas emissions 
ghg_conversion1("hpwh_ref_ghg_natgas",
                "hpwh_intv_ghg_natgas",
                hpwh_ref_data2,
                hpwh_intv_data2,
                fltr.ef_activeyear = 2024, 
                fltr.service_type = "natural gas", 
                fltr.emission_category = "stationary", 
                fltr.emission_scope = "scope 1",
                fltr.service_subcategory1 = "", 
                fltr.service_subcategory2 = "", 
                fltr.country = "global", 
                fltr.subregion = "")

# Calculate scope 2 electricity ghg emissions based on a simulation of grid emissions intensity
ghg_conversion2("hpwh_ref_ghg_elec", "hpwh_intv_ghg_elec", ef_sim_electricity, hpwh_ref_data1, hpwh_intv_data1)

# Calculate scope 3 supply chain ghg emissions based on a simulation of the upstream spend-based emission factor
ghg_conversion3("hpwh_ref_ghg_upstream", "hpwh_intv_ghg_upstream", ef_sim_appliances_upstream, hpwh_ref_capex1, hpwh_intv_capex1)

# Calculate total HPWH GHG emissions
hpwh_ref_ghg_total <- hpwh_ref_ghg_natgas[, -1] + hpwh_ref_ghg_elec[, -1] + hpwh_ref_ghg_upstream[, -1]
hpwh_intv_ghg_total <- hpwh_intv_ghg_natgas[, -1] + hpwh_intv_ghg_elec[, -1] + hpwh_intv_ghg_upstream[, -1]

```

Calculate GHG Mitigation
```{r}
# calculate total reference GHG emissions
ref_ghg_total <- (ashp_ref_ghg_total + weatherize_ref_ghg_total + hpwh_ref_ghg_total)/1000
ref_ghg_total <- ref_ghg_total %>%
    mutate(unit = "mtco2e") %>%  
    select(unit, everything())  

# calculate total intervention GHG emissions
intv_ghg_total <- (ashp_intv_ghg_total + weatherize_intv_ghg_total + hpwh_intv_ghg_total)/1000
intv_ghg_total <- intv_ghg_total %>%
    mutate(unit = "mtco2e") %>%  #
    select(unit, everything())  

# Calculate GHG mitigation
ghg_mitigation <- ref_ghg_total[, -1] - intv_ghg_total[, -1]
ghg_mitigation <- ghg_mitigation %>%
    mutate(unit = "mtco2e") %>%  
    select(unit, everything())  

```

Calculate total costs and co-benefits
```{r}
# Calculate total reference scenario costs
ref_opex_total <- ashp_ref_sumopex + weatherize_ref_sumopex + hpwh_ref_sumopex
ref_capex_total <- ashp_ref_sumcapex + weatherize_ref_sumcapex + hpwh_ref_sumcapex
ref_costs_econ <- ref_opex_total + ref_capex_total

# Calculate total intervention scenario costs
intv_opex_total <- ashp_intv_sumopex + weatherize_intv_sumopex + hpwh_intv_sumopex
intv_capex_total <- ashp_intv_sumcapex + weatherize_intv_sumcapex + hpwh_intv_sumcapex
intv_costs_econ <- intv_opex_total + intv_capex_total

# Social cost of carbon (SCC) is modeled as a normal distribution. It's modeled as an avoided cost, so a positive value is used. 
# mean = $185 sd = $85.7
# The selected value for each simulation is applied consistently over the scenario lifetime.
normal_PDF("coben1_sim", 185, 85.7, units = "USD/mtco2e")

# The function above leads to negative values for the social cost of carbon. In practice, this does not make sense. So negative values are converted to 0 below
coben1_sim[coben1_sim < 0] <- 0

# Calculate social value from Parameter 1 co-benefits 

coben1_value <- ghg_mitigation[, -1] * coben1_sim[, -1]
coben1_value <- coben1_value %>%
    mutate(unit = "social value of avoided carbon (USD)") %>%  
    select(unit, everything()) 
coben_value_trim <- coben1_value[, -1]

# Calculate cashflow
cashflow_investment <- intv_capex_total - ref_capex_total # Assuming capex costs = investment costs
cashflow_opex <- intv_opex_total - ref_opex_total
cashflow_econ <- intv_costs_econ - ref_costs_econ
cashflow_social <- cashflow_econ + coben_value_trim
cashflow_private <- cashflow_econ # This is a space holder for now. This could use the average rebate coverage to calculate cost to city of Flag

```

Excel Exports
```{r}
ashp_outputlist <- (list("ashp_boundary_mgnl" = ashp_boundary_mgnl,
                         "ashp_ref_ann_data1" = ashp_ref_ann_data1,
                         "ashp_ref_ann_data2" = ashp_ref_ann_data2,
                         "ashp_ref_data1" = ashp_ref_data1,
                         "ashp_ref_data2" = ashp_ref_data2,
                         "ashp_ref_ghg_elec" = ashp_ref_ghg_elec,
                         "ashp_ref_ghg_natgas" = ashp_ref_ghg_natgas,
                         "ashp_ref_ghg_total" = ashp_ref_ghg_total,
                         "ashp_ref_ghg_upstream" = ashp_ref_ghg_upstream,
                         "ashp_ref_capex1" = ashp_ref_capex1,
                         "ashp_ref_opex1" = ashp_ref_opex1,
                         "ashp_ref_opex2" = ashp_ref_opex2,
                         "ashp_ref_sumcapex" = ashp_ref_sumcapex,
                         "ashp_ref_sumopex" = ashp_ref_sumopex,
                         "ashp_intv_ann_data1" = ashp_intv_ann_data1,
                         "ashp_intv_ann_data2" = ashp_intv_ann_data2,
                         "ashp_intv_data1" = ashp_intv_data1,
                         "ashp_intv_data2" = ashp_intv_data2,
                         "ashp_intv_ghg_elec" = ashp_intv_ghg_elec,
                         "ashp_intv_ghg_natgas" = ashp_intv_ghg_natgas,
                         "ashp_intv_ghg_total" = ashp_intv_ghg_total,
                         "ashp_intv_ghg_upstream" = ashp_intv_ghg_upstream,
                         "ashp_intv_capex1" = ashp_intv_capex1,
                         "ashp_intv_opex1" = ashp_intv_opex1,
                         "ashp_intv_opex2" = ashp_intv_opex2,
                         "ashp_intv_sumcapex" = ashp_intv_sumcapex,
                         "ashp_intv_sumopex" = ashp_intv_sumopex)
)

weatherize_outputlist <- (list("weatherize_boundary_mgnl" = weatherize_boundary_mgnl,
                         "weatherize_ref_ann_data1" = weatherize_ref_ann_data1,
                         "weatherize_ref_ann_data2" = weatherize_ref_ann_data2,
                         "weatherize_ref_data1" = weatherize_ref_data1,
                         "weatherize_ref_data2" = weatherize_ref_data2,
                         "weatherize_ref_ghg_elec" = weatherize_ref_ghg_elec,
                         "weatherize_ref_ghg_natgas" = weatherize_ref_ghg_natgas,
                         "weatherize_ref_ghg_total" = weatherize_ref_ghg_total,
                         "weatherize_ref_ghg_upstream" = weatherize_ref_ghg_upstream,
                         "weatherize_ref_capex1" = weatherize_ref_capex1,
                         "weatherize_ref_opex1" = weatherize_ref_opex1,
                         "weatherize_ref_opex2" = weatherize_ref_opex2,
                         "weatherize_ref_sumcapex" = weatherize_ref_sumcapex,
                         "weatherize_ref_sumopex" = weatherize_ref_sumopex,
                         "weatherize_intv_ann_data1" = weatherize_intv_ann_data1,
                         "weatherize_intv_ann_data2" = weatherize_intv_ann_data2,
                         "weatherize_intv_data1" = weatherize_intv_data1,
                         "weatherize_intv_data2" = weatherize_intv_data2,
                         "weatherize_intv_ghg_elec" = weatherize_intv_ghg_elec,
                         "weatherize_intv_ghg_natgas" = weatherize_intv_ghg_natgas,
                         "weatherize_intv_ghg_total" = weatherize_intv_ghg_total,
                         "weatherize_intv_ghg_upstream" = weatherize_intv_ghg_upstream,
                         "weatherize_intv_capex1" = weatherize_intv_capex1,
                         "weatherize_intv_opex1" = weatherize_intv_opex1,
                         "weatherize_intv_opex2" = weatherize_intv_opex2,
                         "weatherize_intv_sumcapex" = weatherize_intv_sumcapex,
                         "weatherize_intv_sumopex" = weatherize_intv_sumopex)
)

hpwh_outputlist <- (list("hpwh_boundary_mgnl" = hpwh_boundary_mgnl,
                         "hpwh_ref_ann_data1" = hpwh_ref_ann_data1,
                         "hpwh_ref_ann_data2" = hpwh_ref_ann_data2,
                         "hpwh_ref_data1" = hpwh_ref_data1,
                         "hpwh_ref_data2" = hpwh_ref_data2,
                         "hpwh_ref_ghg_elec" = hpwh_ref_ghg_elec,
                         "hpwh_ref_ghg_natgas" = hpwh_ref_ghg_natgas,
                         "hpwh_ref_ghg_total" = hpwh_ref_ghg_total,
                         "hpwh_ref_ghg_upstream" = hpwh_ref_ghg_upstream,
                         "hpwh_ref_capex1" = hpwh_ref_capex1,
                         "hpwh_ref_opex1" = hpwh_ref_opex1,
                         "hpwh_ref_opex2" = hpwh_ref_opex2,
                         "hpwh_ref_sumcapex" = hpwh_ref_sumcapex,
                         "hpwh_ref_sumopex" = hpwh_ref_sumopex,
                         "hpwh_intv_ann_data1" = hpwh_intv_ann_data1,
                         "hpwh_intv_ann_data2" = hpwh_intv_ann_data2,
                         "hpwh_intv_data1" = hpwh_intv_data1,
                         "hpwh_intv_data2" = hpwh_intv_data2,
                         "hpwh_intv_ghg_elec" = hpwh_intv_ghg_elec,
                         "hpwh_intv_ghg_natgas" = hpwh_intv_ghg_natgas,
                         "hpwh_intv_ghg_total" = hpwh_intv_ghg_total,
                         "hpwh_intv_ghg_upstream" = hpwh_intv_ghg_upstream,
                         "hpwh_intv_capex1" = hpwh_intv_capex1,
                         "hpwh_intv_opex1" = hpwh_intv_opex1,
                         "hpwh_intv_opex2" = hpwh_intv_opex2,
                         "hpwh_intv_sumcapex" = hpwh_intv_sumcapex,
                         "hpwh_intv_sumopex" = hpwh_intv_sumopex)
)

summary_outputlist <- (list("cashflow_investment" = cashflow_investment,
                         "cashflow_opex" = cashflow_opex,
                         "cashflow_private" = cashflow_private,
                         "cashflow_econ" = cashflow_econ,
                         "cashflow_social" = cashflow_social,
                         "coben1_sim" = coben1_sim,
                         "coben1_value" = coben1_value,
                         "ghg_mitigation" = ghg_mitigation,
                         "intv_boundary_mgnl" = intv_boundary_mgnl,
                         "intv_capex_total" = intv_capex_total,
                         "intv_costs_econ" = intv_costs_econ,
                         "intv_ghg_total" = intv_ghg_total,
                         "intv_opex_total" = intv_opex_total,
                         "ref_boundary_mgnl" = ref_boundary_mgnl,
                         "ref_capex_total" = ref_capex_total,
                         "ref_costs_econ" = ref_costs_econ,
                         "ref_ghg_total" = ref_ghg_total,
                         "ref_opex_total" = ref_opex_total)
)

write_xlsx(ashp_outputlist, path = "ashp_output.xlsx")
write_xlsx(weatherize_outputlist, path = "weatherize_output.xlsx")
write_xlsx(hpwh_outputlist, path = "hpwh_output.xlsx")
write_xlsx(summary_outputlist, path = "summary_output.xlsx")

```

Plot: Electricity
```{r}
ref_kwh <- ashp_ref_data1[,-1] + weatherize_ref_data1[,-1] + hpwh_ref_data1[,-1]
ref_kwh <- ref_kwh %>%
    mutate(unit = "kWh") %>%
    select(unit, everything())  

mcs_stats("ref_kWh_stats", ref_kwh, lower_bound = .25, upper_bound = .75)

ref_kWh_stats$year <- as.numeric(ref_kWh_stats$year) # need to convert year column to numeric

# GHG mitigation plot
ggplot(ref_kWh_stats, aes(x = year, y = median)) +
  geom_line(aes(group = 1), color = "black", size = 1) +    # Line for the mean
  geom_ribbon(aes(ymin = lb, ymax = ub, group = 1), fill = "darkgreen", alpha = 0.5) +  # Shaded confidence interval
  labs(title = "ref kwh",
       x = "Year",
       y = "kwh") +
  scale_y_continuous(labels = comma, breaks = scales::breaks_extended(n = 10)) +
  scale_x_continuous(
    breaks = seq(scenario_start, scenario_end, by = 5),
    labels = seq(scenario_start, scenario_end, by = 5)) +
  theme_minimal()

```

Plot: GHG Mitigation
```{r}
mcs_stats("ghg_stats", ghg_mitigation, lower_bound = .25, upper_bound = .75)

ghg_stats$year <- as.numeric(ghg_stats$year) # need to convert year column to numeric

# GHG mitigation plot
ggplot(ghg_stats, aes(x = year, y = ghg_median)) +
  geom_line(aes(group = 1), color = "black", size = 1) +    # Line for the mean
  geom_ribbon(aes(ymin = ghg_lb, ymax = ghg_ub, group = 1), fill = "darkgreen", alpha = 0.5) +  # Shaded confidence interval
  labs(title = "GHG Mitigation",
       x = "Year",
       y = "Metric tons CO2e") +
  scale_y_continuous(labels = comma, breaks = scales::breaks_extended(n = 10)) +
  scale_x_continuous(
    breaks = seq(scenario_start, scenario_end, by = 5),
    labels = seq(scenario_start, scenario_end, by = 5)) +
  theme_minimal()

```

CBA Indicator calcs with discount rate range and Monte Carlo Simulation statistics
```{r}
# Calculates three seperate tables of NPV for private, economic and social costs, using the range of discount rates provided
dr_lb <- 1
dr_ub <- 8

cba_discountrange(dr_min = dr_lb, dr_max = dr_ub, scenario_lifetime = scenario_years)

# Calculates statistics for the Monte Carlo Simulation (lower bound, median, and upper bound)

cba_mcs_stats(lower_bound = .25, upper_bound = .75)

```

Plot: NPV Histograms
```{r}
# Histogram of Economic NPV with 3.5% discount rate
npv_econ <- as.data.frame(npv_econ_dr[,26]) # the 26th row of npv_econ_dr corresponds with a 3% dr. Check npv_stats for column number = dr rate 
npv_econ_lb <- npv_stats[26,5] # Row 26 corresponds to a 3.5 % dr. Column 5, 6, and 7 are the stats for economic npv
npv_econ_median <- npv_stats[26,6]
npv_econ_ub <- npv_stats[26,7]
npv_econ_bin <- (npv_econ_median - npv_econ_lb)/3

ggplot(npv_econ, aes(x = npv_econ[,1])) +
  geom_histogram(binwidth = npv_econ_bin, fill = "darkgrey", color = "black", alpha = 0.7) +
  geom_vline(aes(xintercept = npv_econ_median), color = "darkblue", linetype = "dashed", size = 1) +  # median line
  geom_vline(aes(xintercept = npv_econ_lb), color = "darkgreen", linetype = "dotted", size = 1) +  # lower bound
  geom_vline(aes(xintercept = npv_econ_ub), color = "darkgreen", linetype = "dotted", size = 1) +  # upper bound
  labs(title = "Histogram of Economic NPV at r = 3.5, with median and 25-75% percentiles", x = "npv", y = "Frequency") + 
  scale_x_continuous(labels = comma, breaks = scales::breaks_extended(n = 10)) +
  theme_minimal()

# Histogram of Social NPV with 3% discount rate
npv_social <- as.data.frame(npv_social_dr[,26]) # the 26th row of npv_social_dr corresponds with a 3% dr. Check npv_stats for column number = dr rate 
npv_social_lb <- npv_stats[26,8] # Row 26 corresponds to a 3.5 % dr. Column 5, 6, and 7 are the stats for social npv
npv_social_median <- npv_stats[26,9]
npv_social_ub <- npv_stats[26,10]
npv_social_bin <- (npv_social_median - npv_social_lb)/3

ggplot(npv_social, aes(x = npv_social[,1])) +
  geom_histogram(binwidth = npv_social_bin, fill = "darkgrey", color = "black", alpha = 0.7) +
  geom_vline(aes(xintercept = npv_social_median), color = "darkblue", linetype = "dashed", size = 1) +  # median line
  geom_vline(aes(xintercept = npv_social_lb), color = "darkgreen", linetype = "dotted", size = 1) +  # lower bound
  geom_vline(aes(xintercept = npv_social_ub), color = "darkgreen", linetype = "dotted", size = 1) +  # upper bound
  labs(title = "Histogram of Social NPV at r = 3.5, with median and 25-75% percentiles", x = "npv", y = "Frequency") +
  scale_x_continuous(labels = comma, breaks = scales::breaks_extended(n = 10)) +
  theme_minimal()

```
Plot: NPV Based on discount rate
```{r, eval = FALSE}

npv_stats$dr <- as.numeric(npv_stats$dr) # need to convert discount rate column to numeric

# Econ NPV plot
ggplot(npv_stats, aes(x = dr, y = econ_median)) +
  geom_line(aes(group = 1), color = "black", size = 1) +    # Line for the mean
  geom_ribbon(aes(ymin = econ_lb, ymax = econ_ub, group = 1), fill = "darkgreen", alpha = 0.5) +  # Shaded confidence interval
  labs(title = "NPV based on the discount rate, with 25% - 75% confidence range shaded",
       x = "Discount Rate",
       y = "NPV ($)") +
  scale_y_continuous(labels = comma, breaks = scales::breaks_extended(n = 10)) +
  scale_x_continuous(
    breaks = seq(dr_lb, dr_ub, by = 1),
    labels = seq(dr_lb, dr_ub, by = 1)) +
  theme_minimal()

# Social NPV plot
ggplot(npv_stats, aes(x = dr, y = soc_median)) +
  geom_line(aes(group = 1), color = "black", size = 1) +    # Line for the mean
  geom_ribbon(aes(ymin = soc_lb, ymax = soc_ub, group = 1), fill = "darkblue", alpha = 0.5) +  # Shaded confidence interval
  labs(title = "NPV based on the discount rate, with 25% - 75% confidence range shaded",
       x = "Discount Rate",
       y = "NPV ($)") +
  scale_y_continuous(labels = comma, breaks = scales::breaks_extended(n = 10)) +
  scale_x_continuous(
    breaks = seq(dr_lb, dr_ub, by = 1),
    labels = seq(dr_lb, dr_ub, by = 1)) +
  theme_minimal()
```

Plot: SROI 
```{r}


```


Plot: LCCA 
```{r}


```
