---
title: Monte Carlo Simulation for City of Flagstaff Home Weatherization and Energy Rebate Program
author: Brandon McNamara
project: Dissertation chapter 3
date: March 11, 2025
---

Package Install & Load
```{r setup}
if (!require("dplyr")) install.packages("dplyr")
library(dplyr)
if (!require("readxl")) install.packages("readxl")
library(readxl)
if (!require("writexl")) install.packages("writexl")
library(writexl)
if (!require("ggplot2")) install.packages("ggplot2")
library(ggplot2)
if (!require("scales")) install.packages("scales")
library(scales)
if (!require("data.table")) install.packages("data.table")
library(data.table)
if (!require("gridExtra")) install.packages("gridExtra")
library(gridExtra)
if (!require("patchwork")) install.packages("patchwork")
library(patchwork)
if (!require("tidyr")) install.packages("tidyr")
library(tidyr)
```

Set up global variables
```{r}
# Define the following global variables to be called by parameter functions
run_count <- 10000
scenario_start <- 2025
intv_years <- 10 # The number of years the intervention will be implemented
intv_lifetime <- 15 # The number of years the intervention generates benefits, generally the expected lifetime of equipment
scenario_end  <- scenario_start + intv_years + intv_lifetime - 2
scenario_years <- scenario_end - scenario_start + 1
output_headers <- scenario_start:scenario_end
```

Initiate helper functions
```{r}
source("parameter_functions.r")
```

Create CO2e Emission Factor Library
```{r}
EFL_co2e("ar6")
write_xlsx(EFL, path = "EFLco2e.xlsx") 
```

Emission Factor Simulations
```{r}
# Electricity emission factor
scenario_MCS("ef_sim_electricity", sheet_name = "electricity_ef", units = "kgco2e/kWh") #write a custom function for electricity? with the starting value as a input? And NREL grid region as an input as well. 

# Supply Chain emission factor
continuous_uniform2("ef_sim_appliances_upstream", minvalue = 0.079, maxvalue = 0.161, units = "kgco2e/USD")
continuous_uniform2("ef_sim_weatherize_upstream", minvalue = 0.082, maxvalue = 0.208, units = "kgco2e/USD")

```

Boundary Simulations
```{r}
# Reference boundary represents zero energy efficiency upgrades in each year
static_value("ref_boundary_mgnl", 0, units = "project count")

# Intervention boundary is modeled as a random value between 120 and 160.
# The program is assumed to have funding for 10 years. After that, no more homes are funded. 
discrete_uniform3("intv_boundary_mgnl", minvalue = 120, maxvalue = 160, nyears = intv_years, value2 = 0, units = "project count")

# Project types are randomly assigned for the intervention boundary, with three new dataframes for each project type.
assign_projecttype(intv_boundary_mgnl)

```

Air source heat pumps (ASHP): Activity data, cost, and ghg simulations
```{r}
# Simulate annual electricity consumption for space heating and cooling. Each home in the boundary is simulated individually. The intervention factor is the % energy savings, also simulated for each home in the boundary. 
calc_ad1("ashp_ref_ann_data1",
         "ashp_intv_ann_data1",
         ashp_boundary_mgnl, 
         ref_pdf = rnorm,
         ref_params = list(mean = 4161, sd = 233), # Reference electricity consumption is modeled with a normal distribution
         intv_pdf = runif, 
         intv_params = list(min = 0.31, max = 0.47), # Intervention factor, % savings, is modeled with a uniform distribution  
         units = "kWh")

# Calculate cumulative activity data
cumul_sum2("ashp_ref_data1", ashp_ref_ann_data1)
cumul_sum2("ashp_intv_data1", ashp_intv_ann_data1)

# Simulate annual natural gas consumption for space heating. Each home in the boundary is simulated individually. The intervention factor is the % energy savings, also simulated for each home in the boundary. 
calc_ad1("ashp_ref_ann_data2",
         "ashp_intv_ann_data2",
         ashp_boundary_mgnl, 
         ref_pdf = rnorm,
         ref_params = list(mean = 603, sd = 11.3), # Reference natural gas consumption is modeled with a normal distribution
         intv_pdf = runif, 
         intv_params = list(min = 0.31, max = 0.47), # Intervention factor, % savings,is modeled with a uniform distribution  
         units = "therms")

# Calculate cumulative activity data
cumul_sum2("ashp_ref_data2", ashp_ref_ann_data2)
cumul_sum2("ashp_intv_data2", ashp_intv_ann_data2)

# ASHP Capex is calculated by simulating the installation costs for each home
calculate_capex("ashp_ref_capex1", "ashp_intv_capex1", ref_boundary_mgnl, ashp_boundary_mgnl, capex_min = 2100, capex_max = 6000, units = "USD")

# Sum total Capex
ashp_ref_sumcapex <- ashp_ref_capex1[,-1]
ashp_intv_sumcapex <- ashp_intv_capex1[,-1]

# ASHP Opex is calculated for electricity and natural gas separately, based on current energy prices
calculate_opex("ashp_ref_opex1", "ashp_intv_opex1", ashp_ref_data1, ashp_intv_data1, opex = 0.1542, units = "USD")
calculate_opex("ashp_ref_opex2", "ashp_intv_opex2", ashp_ref_data2, ashp_intv_data2, opex = 0.3434, units = "therms")

# Sum total opex
ashp_ref_sumopex <- ashp_ref_opex1[,-1] + ashp_ref_opex2[,-1]
ashp_intv_sumopex <- ashp_intv_opex1[,-1] + ashp_intv_opex2[,-1]

# Calculate scope 1 natural gas emissions 
ghg_conversion1("ashp_ref_ghg_natgas",
                "ashp_intv_ghg_natgas",
                ashp_ref_data2,
                ashp_intv_data2,
                fltr.ef_activeyear = 2024, 
                fltr.service_type = "natural gas", 
                fltr.emission_category = "stationary", 
                fltr.emission_scope = "scope 1",
                fltr.service_subcategory1 = "", 
                fltr.service_subcategory2 = "", 
                fltr.country = "global", 
                fltr.subregion = "")

# Calculate scope 2 electricity ghg emissions based on a simulation of grid emissions intensity
ghg_conversion2("ashp_ref_ghg_elec", "ashp_intv_ghg_elec", ef_sim_electricity, ashp_ref_data1, ashp_intv_data1)

# Calculate scope 3 supply chain ghg emissions based on a simulation of the upstream spend-based emission factor
ghg_conversion3("ashp_ref_ghg_upstream", "ashp_intv_ghg_upstream", ef_sim_appliances_upstream, ashp_ref_capex1, ashp_intv_capex1)

# Calculate total ashp GHG emissions
ashp_ref_ghg_total <- ashp_ref_ghg_natgas[, -1] + ashp_ref_ghg_elec[, -1] + ashp_ref_ghg_upstream[, -1]
ashp_intv_ghg_total <- ashp_intv_ghg_natgas[, -1] + ashp_intv_ghg_elec[, -1] + ashp_intv_ghg_upstream[, -1]

```

Weatherization: Activity data, cost, and GHG simulations
```{r}
# Simulate annual electricity consumption for space heating and cooling. Each home in the boundary is simulated individually. The intervention factor is the % energy savings, also simulated for each home in the boundary. 
calc_ad1("weatherize_ref_ann_data1",
         "weatherize_intv_ann_data1",
         weatherize_boundary_mgnl, 
         ref_pdf = rnorm,
         ref_params = list(mean = 4161, sd = 233), # Reference electricity consumption is modeled with a normal distribution
         intv_pdf = runif, 
         intv_params = list(min = 0.10, max = 0.20), # Intervention factor, % savings, is modeled with a uniform distribution  
         units = "kWh")

# Calculate cumulative activity data
cumul_sum2("weatherize_ref_data1", weatherize_ref_ann_data1)
cumul_sum2("weatherize_intv_data1", weatherize_intv_ann_data1)

# Simulate annual natural gas consumption for space heating. Each home in the boundary is simulated individually. The intervention factor is the % energy savings, also simulated for each home in the boundary. 
calc_ad1("weatherize_ref_ann_data2",
         "weatherize_intv_ann_data2",
         weatherize_boundary_mgnl, 
         ref_pdf = rnorm,
         ref_params = list(mean = 603, sd = 11.3), # Reference natural gas consumption is modeled with a normal distribution
         intv_pdf = runif, 
         intv_params = list(min = 0.10, max = 0.20), # Intervention factor, % savings,is modeled with a uniform distribution  
         units = "therms")

# Calculate cumulative activity data
cumul_sum2("weatherize_ref_data2", weatherize_ref_ann_data2)
cumul_sum2("weatherize_intv_data2", weatherize_intv_ann_data2)

# Weatherization Capex is calculated by simulating the installation costs for each home
calculate_capex("weatherize_ref_capex1", "weatherize_intv_capex1", ref_boundary_mgnl, weatherize_boundary_mgnl, capex_min = 1879, capex_max = 2962, units = "USD")

# Sum total Capex
weatherize_ref_sumcapex <- weatherize_ref_capex1[,-1]
weatherize_intv_sumcapex <- weatherize_intv_capex1[,-1]

# Weatherization Opex is calculated for electricity and natural gas separately, based on current energy prices
calculate_opex("weatherize_ref_opex1", "weatherize_intv_opex1", weatherize_ref_data1, weatherize_intv_data1, opex = 0.1542, units = "USD")
calculate_opex("weatherize_ref_opex2", "weatherize_intv_opex2", weatherize_ref_data2, weatherize_intv_data2, opex = 0.3434, units = "therms")

# Sum total Opex
weatherize_ref_sumopex <- weatherize_ref_opex1[,-1] + weatherize_ref_opex2[,-1]
weatherize_intv_sumopex <- weatherize_intv_opex1[,-1] + weatherize_intv_opex2[,-1]

# Calculate scope 1 natural gas emissions 
ghg_conversion1("weatherize_ref_ghg_natgas",
                "weatherize_intv_ghg_natgas",
                weatherize_ref_data2,
                weatherize_intv_data2,
                fltr.ef_activeyear = 2024, 
                fltr.service_type = "natural gas", 
                fltr.emission_category = "stationary", 
                fltr.emission_scope = "scope 1",
                fltr.service_subcategory1 = "", 
                fltr.service_subcategory2 = "", 
                fltr.country = "global", 
                fltr.subregion = "")

# Calculate scope 2 electricity ghg emissions based on a simulation of grid emissions intensity
ghg_conversion2("weatherize_ref_ghg_elec", "weatherize_intv_ghg_elec", ef_sim_electricity, weatherize_ref_data1, weatherize_intv_data1)

# Calculate scope 3 supply chain ghg emissions based on a simulation of the upstream spend-based emission factor
ghg_conversion3("weatherize_ref_ghg_upstream", "weatherize_intv_ghg_upstream", ef_sim_weatherize_upstream, weatherize_ref_capex1, weatherize_intv_capex1)

# Calculate total weatherize GHG emissions
weatherize_ref_ghg_total <- weatherize_ref_ghg_natgas[, -1] + weatherize_ref_ghg_elec[, -1] + weatherize_ref_ghg_upstream[, -1]
weatherize_intv_ghg_total <- weatherize_intv_ghg_natgas[, -1] + weatherize_intv_ghg_elec[, -1] + weatherize_intv_ghg_upstream[, -1]

```

Heat Pump Water Heater (HPWH): Activity data, cost, and GHG simulations
```{r}
# Simulate annual electricity consumption for water heating. Each home in the boundary is simulated individually. The intervention factor is the % energy savings, also simulated for each home in the boundary. 
calc_ad1("hpwh_ref_ann_data1",
         "hpwh_intv_ann_data1",
         hpwh_boundary_mgnl, 
         ref_pdf = rnorm,
         ref_params = list(mean = 2969, sd = 108.1), # Reference electricity consumption is modeled with a normal distribution
         intv_pdf = runif, 
         intv_params = list(min = 0.30, max = 0.70), # Intervention factor, % savings, is modeled with a uniform distribution  
         units = "kWh")

# Calculate cumulative activity data
cumul_sum2("hpwh_ref_data1", hpwh_ref_ann_data1)
cumul_sum2("hpwh_intv_data1", hpwh_intv_ann_data1)

# Simulate annual natural gas consumption for space heating. Each home in the boundary is simulated individually. The intervention factor is the % energy savings, also simulated for each home in the boundary. 
calc_ad1("hpwh_ref_ann_data2",
         "hpwh_intv_ann_data2",
         hpwh_boundary_mgnl, 
         ref_pdf = rnorm,
         ref_params = list(mean = 237, sd = 4), # Reference natural gas consumption is modeled with a normal distribution
         intv_pdf = runif, 
         intv_params = list(min = 0.30, max = 0.70), # Intervention factor, % savings,is modeled with a uniform distribution  
         units = "therms")

# Calculate cumulative activity data
cumul_sum2("hpwh_ref_data2", hpwh_ref_ann_data2)
cumul_sum2("hpwh_intv_data2", hpwh_intv_ann_data2)

# HPWH Capex is calculated by simulating the installation costs for each home
calculate_capex("hpwh_ref_capex1", "hpwh_intv_capex1", ref_boundary_mgnl, hpwh_boundary_mgnl, capex_min = 1300, capex_max = 2900, units = "USD")

# Sum total Capex
hpwh_ref_sumcapex <- hpwh_ref_capex1[,-1]
hpwh_intv_sumcapex <- hpwh_intv_capex1[,-1]

# HPWH Opex is calculated for electricity and natural gas separately, based on current energy prices
calculate_opex("hpwh_ref_opex1", "hpwh_intv_opex1", hpwh_ref_data1, hpwh_intv_data1, opex = 0.1542, units = "USD")
calculate_opex("hpwh_ref_opex2", "hpwh_intv_opex2", hpwh_ref_data2, hpwh_intv_data2, opex = 0.3434, units = "therms")

# Sum total Opex
hpwh_ref_sumopex <- hpwh_ref_opex1[,-1] + hpwh_ref_opex2[,-1]
hpwh_intv_sumopex <- hpwh_intv_opex1[,-1] + hpwh_intv_opex2[,-1]

# Calculate scope 1 natural gas emissions 
ghg_conversion1("hpwh_ref_ghg_natgas",
                "hpwh_intv_ghg_natgas",
                hpwh_ref_data2,
                hpwh_intv_data2,
                fltr.ef_activeyear = 2024, 
                fltr.service_type = "natural gas", 
                fltr.emission_category = "stationary", 
                fltr.emission_scope = "scope 1",
                fltr.service_subcategory1 = "", 
                fltr.service_subcategory2 = "", 
                fltr.country = "global", 
                fltr.subregion = "")

# Calculate scope 2 electricity ghg emissions based on a simulation of grid emissions intensity
ghg_conversion2("hpwh_ref_ghg_elec", "hpwh_intv_ghg_elec", ef_sim_electricity, hpwh_ref_data1, hpwh_intv_data1)

# Calculate scope 3 supply chain ghg emissions based on a simulation of the upstream spend-based emission factor
ghg_conversion3("hpwh_ref_ghg_upstream", "hpwh_intv_ghg_upstream", ef_sim_appliances_upstream, hpwh_ref_capex1, hpwh_intv_capex1)

# Calculate total HPWH GHG emissions
hpwh_ref_ghg_total <- hpwh_ref_ghg_natgas[, -1] + hpwh_ref_ghg_elec[, -1] + hpwh_ref_ghg_upstream[, -1]
hpwh_intv_ghg_total <- hpwh_intv_ghg_natgas[, -1] + hpwh_intv_ghg_elec[, -1] + hpwh_intv_ghg_upstream[, -1]

```

Calculate GHG Mitigation
```{r}
# calculate total reference GHG emissions
ref_ghg_total <- (ashp_ref_ghg_total + weatherize_ref_ghg_total + hpwh_ref_ghg_total)/1000
ref_ghg_total <- ref_ghg_total %>%
    mutate(unit = "mtco2e") %>%  
    select(unit, everything())  

# calculate total intervention GHG emissions
intv_ghg_total <- (ashp_intv_ghg_total + weatherize_intv_ghg_total + hpwh_intv_ghg_total)/1000
intv_ghg_total <- intv_ghg_total %>%
    mutate(unit = "mtco2e") %>%  #
    select(unit, everything())  

# Calculate GHG mitigation
ghg_mitigation <- ref_ghg_total[, -1] - intv_ghg_total[, -1]
ghg_mitigation <- ghg_mitigation %>%
    mutate(unit = "mtco2e") %>%  
    select(unit, everything())  

```

Calculate total costs and co-benefits
```{r}
# Calculate total reference scenario costs
ref_opex_total <- ashp_ref_sumopex + weatherize_ref_sumopex + hpwh_ref_sumopex
ref_capex_total <- ashp_ref_sumcapex + weatherize_ref_sumcapex + hpwh_ref_sumcapex
ref_costs_econ <- ref_opex_total + ref_capex_total

# Calculate total intervention scenario costs
intv_opex_total <- ashp_intv_sumopex + weatherize_intv_sumopex + hpwh_intv_sumopex
intv_capex_total <- ashp_intv_sumcapex + weatherize_intv_sumcapex + hpwh_intv_sumcapex
intv_costs_econ <- intv_opex_total + intv_capex_total

# Social cost of carbon (SCC) is modeled as a normal distribution. It's modeled as an avoided cost, so a positive value is used. 
# mean = $185 sd = $85.7
# The selected value for each simulation is applied consistently over the scenario lifetime.
normal_PDF("coben1_sim", 185, 85.7, units = "USD/mtco2e")

# The function above leads to negative values for the social cost of carbon. In practice, this does not make sense. So negative values are converted to 0 below
coben1_sim[coben1_sim < 0] <- 0

# Calculate social value from Parameter 1 co-benefits 

coben1_value <- ghg_mitigation[, -1] * coben1_sim[, -1]
coben1_value <- coben1_value %>%
    mutate(unit = "social value of avoided carbon (USD)") %>%  
    select(unit, everything()) 
coben_value_trim <- coben1_value[, -1]

# Calculate cashflow
cashflow_investment <- intv_capex_total - ref_capex_total # Assuming capex costs = investment costs
cashflow_opex <- intv_opex_total - ref_opex_total
cashflow_econ <- intv_costs_econ - ref_costs_econ
cashflow_social <- cashflow_econ + coben_value_trim
cashflow_private <- cashflow_econ # This is a space holder for now. This could use the average rebate coverage to calculate cost to city of Flag

```

Excel Exports
```{r}
ashp_outputlist <- (list("ashp_boundary_mgnl" = ashp_boundary_mgnl,
                         "ashp_ref_ann_data1" = ashp_ref_ann_data1,
                         "ashp_ref_ann_data2" = ashp_ref_ann_data2,
                         "ashp_ref_data1" = ashp_ref_data1,
                         "ashp_ref_data2" = ashp_ref_data2,
                         "ashp_ref_ghg_elec" = ashp_ref_ghg_elec,
                         "ashp_ref_ghg_natgas" = ashp_ref_ghg_natgas,
                         "ashp_ref_ghg_total" = ashp_ref_ghg_total,
                         "ashp_ref_ghg_upstream" = ashp_ref_ghg_upstream,
                         "ashp_ref_capex1" = ashp_ref_capex1,
                         "ashp_ref_opex1" = ashp_ref_opex1,
                         "ashp_ref_opex2" = ashp_ref_opex2,
                         "ashp_ref_sumcapex" = ashp_ref_sumcapex,
                         "ashp_ref_sumopex" = ashp_ref_sumopex,
                         "ashp_intv_ann_data1" = ashp_intv_ann_data1,
                         "ashp_intv_ann_data2" = ashp_intv_ann_data2,
                         "ashp_intv_data1" = ashp_intv_data1,
                         "ashp_intv_data2" = ashp_intv_data2,
                         "ashp_intv_ghg_elec" = ashp_intv_ghg_elec,
                         "ashp_intv_ghg_natgas" = ashp_intv_ghg_natgas,
                         "ashp_intv_ghg_total" = ashp_intv_ghg_total,
                         "ashp_intv_ghg_upstream" = ashp_intv_ghg_upstream,
                         "ashp_intv_capex1" = ashp_intv_capex1,
                         "ashp_intv_opex1" = ashp_intv_opex1,
                         "ashp_intv_opex2" = ashp_intv_opex2,
                         "ashp_intv_sumcapex" = ashp_intv_sumcapex,
                         "ashp_intv_sumopex" = ashp_intv_sumopex)
)

weatherize_outputlist <- (list("weatherize_boundary_mgnl" = weatherize_boundary_mgnl,
                         "weatherize_ref_ann_data1" = weatherize_ref_ann_data1,
                         "weatherize_ref_ann_data2" = weatherize_ref_ann_data2,
                         "weatherize_ref_data1" = weatherize_ref_data1,
                         "weatherize_ref_data2" = weatherize_ref_data2,
                         "weatherize_ref_ghg_elec" = weatherize_ref_ghg_elec,
                         "weatherize_ref_ghg_natgas" = weatherize_ref_ghg_natgas,
                         "weatherize_ref_ghg_total" = weatherize_ref_ghg_total,
                         "weatherize_ref_ghg_upstream" = weatherize_ref_ghg_upstream,
                         "weatherize_ref_capex1" = weatherize_ref_capex1,
                         "weatherize_ref_opex1" = weatherize_ref_opex1,
                         "weatherize_ref_opex2" = weatherize_ref_opex2,
                         "weatherize_ref_sumcapex" = weatherize_ref_sumcapex,
                         "weatherize_ref_sumopex" = weatherize_ref_sumopex,
                         "weatherize_intv_ann_data1" = weatherize_intv_ann_data1,
                         "weatherize_intv_ann_data2" = weatherize_intv_ann_data2,
                         "weatherize_intv_data1" = weatherize_intv_data1,
                         "weatherize_intv_data2" = weatherize_intv_data2,
                         "weatherize_intv_ghg_elec" = weatherize_intv_ghg_elec,
                         "weatherize_intv_ghg_natgas" = weatherize_intv_ghg_natgas,
                         "weatherize_intv_ghg_total" = weatherize_intv_ghg_total,
                         "weatherize_intv_ghg_upstream" = weatherize_intv_ghg_upstream,
                         "weatherize_intv_capex1" = weatherize_intv_capex1,
                         "weatherize_intv_opex1" = weatherize_intv_opex1,
                         "weatherize_intv_opex2" = weatherize_intv_opex2,
                         "weatherize_intv_sumcapex" = weatherize_intv_sumcapex,
                         "weatherize_intv_sumopex" = weatherize_intv_sumopex)
)

hpwh_outputlist <- (list("hpwh_boundary_mgnl" = hpwh_boundary_mgnl,
                         "hpwh_ref_ann_data1" = hpwh_ref_ann_data1,
                         "hpwh_ref_ann_data2" = hpwh_ref_ann_data2,
                         "hpwh_ref_data1" = hpwh_ref_data1,
                         "hpwh_ref_data2" = hpwh_ref_data2,
                         "hpwh_ref_ghg_elec" = hpwh_ref_ghg_elec,
                         "hpwh_ref_ghg_natgas" = hpwh_ref_ghg_natgas,
                         "hpwh_ref_ghg_total" = hpwh_ref_ghg_total,
                         "hpwh_ref_ghg_upstream" = hpwh_ref_ghg_upstream,
                         "hpwh_ref_capex1" = hpwh_ref_capex1,
                         "hpwh_ref_opex1" = hpwh_ref_opex1,
                         "hpwh_ref_opex2" = hpwh_ref_opex2,
                         "hpwh_ref_sumcapex" = hpwh_ref_sumcapex,
                         "hpwh_ref_sumopex" = hpwh_ref_sumopex,
                         "hpwh_intv_ann_data1" = hpwh_intv_ann_data1,
                         "hpwh_intv_ann_data2" = hpwh_intv_ann_data2,
                         "hpwh_intv_data1" = hpwh_intv_data1,
                         "hpwh_intv_data2" = hpwh_intv_data2,
                         "hpwh_intv_ghg_elec" = hpwh_intv_ghg_elec,
                         "hpwh_intv_ghg_natgas" = hpwh_intv_ghg_natgas,
                         "hpwh_intv_ghg_total" = hpwh_intv_ghg_total,
                         "hpwh_intv_ghg_upstream" = hpwh_intv_ghg_upstream,
                         "hpwh_intv_capex1" = hpwh_intv_capex1,
                         "hpwh_intv_opex1" = hpwh_intv_opex1,
                         "hpwh_intv_opex2" = hpwh_intv_opex2,
                         "hpwh_intv_sumcapex" = hpwh_intv_sumcapex,
                         "hpwh_intv_sumopex" = hpwh_intv_sumopex)
)

summary_outputlist <- (list("cashflow_investment" = cashflow_investment,
                         "cashflow_opex" = cashflow_opex,
                         "cashflow_private" = cashflow_private,
                         "cashflow_econ" = cashflow_econ,
                         "cashflow_social" = cashflow_social,
                         "coben1_sim" = coben1_sim,
                         "coben1_value" = coben1_value,
                         "ghg_mitigation" = ghg_mitigation,
                         "intv_boundary_mgnl" = intv_boundary_mgnl,
                         "intv_capex_total" = intv_capex_total,
                         "intv_costs_econ" = intv_costs_econ,
                         "intv_ghg_total" = intv_ghg_total,
                         "intv_opex_total" = intv_opex_total,
                         "ref_boundary_mgnl" = ref_boundary_mgnl,
                         "ref_capex_total" = ref_capex_total,
                         "ref_costs_econ" = ref_costs_econ,
                         "ref_ghg_total" = ref_ghg_total,
                         "ref_opex_total" = ref_opex_total)
)

write_xlsx(ashp_outputlist, path = "ashp_output.xlsx")
write_xlsx(weatherize_outputlist, path = "weatherize_output.xlsx")
write_xlsx(hpwh_outputlist, path = "hpwh_output.xlsx")
write_xlsx(summary_outputlist, path = "summary_output.xlsx")

```

Figure 1: Energy use in reference & intervention scenarios
```{r}
# Electricity

# Create dataframe with total mwh for each scenario
ref_mwh <- ashp_ref_data1[,-1] + weatherize_ref_data1[,-1] + hpwh_ref_data1[,-1]
ref_mwh <- ref_mwh / 1000
intv_mwh <- ashp_intv_data1[,-1] + weatherize_intv_data1[,-1] + hpwh_intv_data1[,-1]
intv_mwh <- intv_mwh / 1000

mcs_stats_cumul("ref_mwh_stats", ref_mwh, lower_bound_out = .05, lower_bound_in = .25, upper_bound_in = .75, upper_bound_out = .95)
mcs_stats_cumul("intv_mwh_stats", intv_mwh, lower_bound_out = .05, lower_bound_in = .25, upper_bound_in = .75, upper_bound_out = .95)

# Convert year column to numeric
ref_mwh_stats$year <- as.numeric(ref_mwh_stats$year) 
intv_mwh_stats$year <- as.numeric(intv_mwh_stats$year) 

# Add an identifier column to each dataframe
ref_mwh_stats$Scenario <- "Reference Scenario"
intv_mwh_stats$Scenario <- "Intervention Scenario"

mwhplot <- bind_rows(ref_mwh_stats, intv_mwh_stats)

# mwh comparison plot
p_mwh <- ggplot(mwhplot, aes(x = year, y = median, color = Scenario, fill = Scenario)) +
  geom_line(size = .5) + 
  geom_ribbon(aes(ymin = lb1, ymax = ub1), alpha = 0.25, color = NA) +
  geom_ribbon(aes(ymin = lb2, ymax = ub2), alpha = 0.4, color = NA) +
  labs(title = "a",
       x = "Year",
       y = "Megawatt-hours") +
  scale_y_continuous(labels = comma, breaks = scales::breaks_extended(n = 6)) +
  scale_x_continuous(
    breaks = seq(scenario_start, scenario_end, by = 5),
    labels = seq(scenario_start, scenario_end, by = 5)) +
  theme_minimal() +
  scale_color_manual(values = c("Reference Scenario" = "darkblue", "Intervention Scenario" = "darkgreen")) +
  scale_fill_manual(values = c("Reference Scenario" = "darkblue", "Intervention Scenario" = "darkgreen")) +
  theme(text = element_text(size = 8, color = "black"),
        plot.title = element_text(face = "bold", size = 12, color = "black"),
        axis.text = element_text(size = 8, color = "black"),
        axis.title = element_text(size = 8, color = "black"),
        legend.text = element_text(size = 8, color = "black"),
        axis.ticks = element_line(color = "black", size = .5),
        axis.line = element_line(color = "black", size = .5),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        aspect.ratio = 1)

# Natural Gas

# Create dataframe with total mmbtu for each scenario
ref_mmbtu <- ashp_ref_data2[,-1] + weatherize_ref_data2[,-1] + hpwh_ref_data2[,-1]
ref_mmbtu <- ref_mmbtu / 10
intv_mmbtu <- ashp_intv_data2[,-1] + weatherize_intv_data2[,-1] + hpwh_intv_data2[,-1]
intv_mmbtu <- intv_mmbtu / 10

mcs_stats_cumul("ref_mmbtu_stats", ref_mmbtu, lower_bound_out = .05, lower_bound_in = .25, upper_bound_in = .75, upper_bound_out = .95)
mcs_stats_cumul("intv_mmbtu_stats", intv_mmbtu, lower_bound_out = .05, lower_bound_in = .25, upper_bound_in = .75, upper_bound_out = .95)

# Convert year column to numeric
ref_mmbtu_stats$year <- as.numeric(ref_mmbtu_stats$year) 
intv_mmbtu_stats$year <- as.numeric(intv_mmbtu_stats$year) 

# Add an identifier column to each dataframe
ref_mmbtu_stats$Scenario <- "Reference Scenario"
intv_mmbtu_stats$Scenario <- "Intervention Scenario"

mmbtuplot <- bind_rows(ref_mmbtu_stats, intv_mmbtu_stats)

# mmbtu comparison plot
p_mmbtu <- ggplot(mmbtuplot, aes(x = year, y = median, color = Scenario, fill = Scenario)) +
  geom_line(size = .5) + 
  geom_ribbon(aes(ymin = lb1, ymax = ub1), alpha = 0.25, color = NA) +
  geom_ribbon(aes(ymin = lb2, ymax = ub2), alpha = 0.4, color = NA) +
  labs(title = "b",
       x = "Year",
       y = "Million BTUs") +
  scale_y_continuous(labels = comma, breaks = scales::breaks_extended(n = 6)) +
  scale_x_continuous(
    breaks = seq(scenario_start, scenario_end, by = 5),
    labels = seq(scenario_start, scenario_end, by = 5)) +
  theme_minimal() +
  scale_color_manual(values = c("Reference Scenario" = "darkblue", "Intervention Scenario" = "darkgreen")) +
  scale_fill_manual(values = c("Reference Scenario" = "darkblue", "Intervention Scenario" = "darkgreen")) +
  theme(text = element_text(size = 8, color = "black"),
        plot.title = element_text(face = "bold", size = 12, color = "black"),
        axis.text = element_text(size = 8, color = "black"),
        axis.title = element_text(size = 8, color = "black"),
        legend.text = element_text(size = 8, color = "black"),
        axis.ticks = element_line(color = "black", size = .5),
        axis.line = element_line(color = "black", size = .5),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        aspect.ratio = 1)

p1 <- p_mwh + p_mmbtu + plot_layout(guides = 'collect')

ggsave("figures/fig1.png", plot = p1, width = 6.5, height = 2.5, dpi = 300)

```

Figure 2: Capex and Opex
```{r}
# Capex & opex cashflow

quick_discount("capex3.5", cashflow_investment, dr = 3.5)
quick_discount("opex3.5", cashflow_opex, dr = 3.5)

mcs_stats_ann("capex3.5_stats", capex3.5, lower_bound_out = .05, lower_bound_in = .25, upper_bound_in = .75, upper_bound_out = .95)
mcs_stats_ann("opex3.5_stats", opex3.5, lower_bound_out = .05, lower_bound_in = .25, upper_bound_in = .75, upper_bound_out = .95)

# Convert year column to numeric
capex3.5_stats$year <- as.numeric(capex3.5_stats$year) 
opex3.5_stats$year <- as.numeric(opex3.5_stats$year) 

# capex plot
p_capex <- ggplot(capex3.5_stats, aes(x = year, y = median)) +
  geom_line(aes(group = 1), color = "deepskyblue4", size = .5) +   
  geom_ribbon(aes(ymin = lb1, ymax = ub1, group = 1), fill = "deepskyblue4", alpha = 0.25) + 
  geom_ribbon(aes(ymin = lb2, ymax = ub2, group = 1), fill = "deepskyblue4", alpha = 0.4) +  
  labs(title = "a",
       x = "Year",
       y = "Annual Capex\n(present value USD)") +
  scale_y_continuous(labels = comma, breaks = scales::breaks_extended(n = 6)) +
  scale_x_continuous(
    breaks = seq(scenario_start, scenario_end, by = 10),
    labels = seq(scenario_start, scenario_end, by = 10)) +
  theme_minimal() +
  theme(text = element_text(size = 8, color = "black"),
        plot.title = element_text(face = "bold", size = 12, color = "black"),
        axis.text = element_text(size = 8, color = "black"),
        axis.title = element_text(size = 8, color = "black"),
        legend.text = element_text(size = 8, color = "black"),
        axis.ticks = element_line(color = "black", size = .5),
        axis.line = element_line(color = "black", size = .5),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        aspect.ratio = 1)

# opex plot
p_opex <- ggplot(opex3.5_stats, aes(x = year, y = median)) +
  geom_line(aes(group = 1), color = "palegreen4", size = .5) +   
  geom_ribbon(aes(ymin = lb1, ymax = ub1, group = 1), fill = "palegreen4", alpha = 0.25) + 
  geom_ribbon(aes(ymin = lb2, ymax = ub2, group = 1), fill = "palegreen4", alpha = 0.4) +  
  labs(title = "b",
       x = "Year",
       y = "Annual Opex\n(present value USD)") +
  scale_y_continuous(labels = comma, breaks = scales::breaks_extended(n = 6)) +
  scale_x_continuous(
    breaks = seq(scenario_start, scenario_end, by = 10),
    labels = seq(scenario_start, scenario_end, by = 10)) +
  theme_minimal() +
  theme(text = element_text(size = 8, color = "black"),
        plot.title = element_text(face = "bold", size = 12, color = "black"),
        axis.text = element_text(size = 8, color = "black"),
        axis.title = element_text(size = 8, color = "black"),
        legend.text = element_text(size = 8, color = "black"),
        axis.ticks = element_line(color = "black", size = .5),
        axis.line = element_line(color = "black", size = .5),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        aspect.ratio = 1)

# Economic cashflow plot 

quick_discount("cashflow_econ3.5", cashflow_econ, dr = 3.5)
mcs_stats_cumul("cashflow_stats", cashflow_econ3.5, lower_bound_out = .05, lower_bound_in = .25, upper_bound_in = .75, upper_bound_out = .95)

# Convert year column to numeric
cashflow_stats$year <- as.numeric(cashflow_stats$year) 

# Plot
p_cashflow <- ggplot(cashflow_stats, aes(x = year, y = median)) +
  geom_line(aes(group = 1), color = "sienna", size = .5) +   
  geom_ribbon(aes(ymin = lb1, ymax = ub1, group = 1), fill = "sienna", alpha = 0.25) + 
  geom_ribbon(aes(ymin = lb2, ymax = ub2, group = 1), fill = "sienna", alpha = 0.4) +  
  labs(title = "c",
       x = "Year",
       y = "Cumulative Cashflow\n(present value USD)") +
  scale_y_continuous(labels = comma, breaks = scales::breaks_extended(n = 6)) +
  scale_x_continuous(
    breaks = seq(scenario_start, scenario_end, by = 10),
    labels = seq(scenario_start, scenario_end, by = 10)) +
  theme_minimal() +
  theme(text = element_text(size = 8, color = "black"),
        plot.title = element_text(face = "bold", size = 12, color = "black"),
        axis.text = element_text(size = 8, color = "black"),
        axis.title = element_text(size = 8, color = "black"),
        legend.text = element_text(size = 8, color = "black"),
        axis.ticks = element_line(color = "black", size = .5),
        axis.line = element_line(color = "black", size = .5),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        aspect.ratio = 1)

p2 <- p_capex + p_opex + p_cashflow

ggsave("figures/fig2.png", plot = p2, width = 6.5, height = 2, dpi = 300)

```


Figure 3: GHG Mitigation
```{r}
# GHG mitigation histogram
cumul_sum1("ghg_cumul", ghg_mitigation)
mcs_stats_cumul("ghg_stats", ghg_mitigation[, -1], lower_bound_out = .05, lower_bound_in = .25, upper_bound_in = .75, upper_bound_out = .95)
total_ghg_mit <- as.data.frame(ghg_cumul[,(scenario_years + 1)]) 
ghg_lb1 <- ghg_stats[scenario_years,2]
ghg_lb2 <- ghg_stats[scenario_years,3] 
ghg_median <- ghg_stats[scenario_years,4]
ghg_ub2 <- ghg_stats[scenario_years,5]
ghg_ub1 <- ghg_stats[scenario_years,6]
ghg_bin <- (ghg_median - ghg_lb2)/3

p3 <- ggplot(total_ghg_mit, aes(x = total_ghg_mit[,1])) +
  geom_histogram(binwidth = ghg_bin, fill = "darkgrey", color = "black", alpha = 0.7) +
  geom_vline(aes(xintercept = ghg_median, color = "Median"), linetype = "dashed", size = 1) + 
  geom_vline(aes(xintercept = ghg_lb1, color = "5%-95% Quantile Range"), linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = ghg_lb2, color = "25%-75% Quantile Range"), linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = ghg_ub2), color = "orange3", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = ghg_ub1), color = "darkgreen", linetype = "dashed", size = 1) +
  labs(title = NULL, x = "tco2e", y = "Frequency") + 
  scale_x_continuous(labels = comma, breaks = scales::breaks_extended(n = 10)) +
  theme_minimal() + 
  theme(text = element_text(size = 8, color = "black"),
        plot.title = element_text(face = "bold", size = 12, color = "black"),
        axis.text = element_text(size = 8, color = "black"),
        axis.title = element_text(size = 8, color = "black"),
        legend.text = element_text(size = 8, color = "black"),
        axis.ticks = element_line(color = "black", size = .5),
        axis.line = element_line(color = "black", size = .5),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        aspect.ratio = .5,
        legend.position = "right",  
        legend.title = element_blank()) + 
  scale_color_manual(name = NULL, 
                     values = c("Median" = "darkblue",
                                "25%-75% Quantile Range" = "orange3",
                                "5%-95% Quantile Range" = "darkgreen"),
                     breaks = c("Median", "25%-75% Quantile Range", "5%-95% Quantile Range"),
                     guide = guide_legend(override.aes = list(linetype = 1)))

ggsave("figures/fig3.png", plot = p3, width = 6.5, height = 3, dpi = 300)

```

Figure 4: Electricity EF
```{r}
# Electricity scenarios
elec_ef_scenarios <- read_excel("scenarios.xlsx", sheet = "electricity_ef")
fig4_trim <- cbind(elec_ef_scenarios[, 3], elec_ef_scenarios[, 9:34])
fig4_data <- fig4_trim %>%
  pivot_longer(cols = starts_with("20"), 
               names_to = "year", 
               values_to = "value")
fig4_data$year <- as.numeric(fig4_data$year)
colnames(fig4_data)[1] <- "scenario"

p4_scenario_order <- c("Mid-case", 
                    "Low Renewable Energy Costs", 
                    "High Renewable Energy Costs", 
                    "Low Natural Gas Prices", 
                    "High Natural Gas Prices", 
                    "High Demand Growth", 
                    "95% Decarbonization by 2050", 
                    "100% Decarbonization by 2035")

p4_colors <- c(
  "Mid-case" = "#1F77B4",  # Blue
  "Low Renewable Energy Costs" = "#FF7F0E",  # Orange
  "High Renewable Energy Costs" = "#2CA02C",  # Green
  "Low Natural Gas Prices" = "#D62728",  # Red
  "High Natural Gas Prices" = "#9467BD",  # Purple
  "High Demand Growth" = "#8C564B",  # Brown
  "95% Decarbonization by 2050" = "#E377C2",  # Light Purple
  "100% Decarbonization by 2035" = "#7F7F7F"  # Gray
)

p4 <- ggplot(fig4_data, aes(x = year, y = value, color = scenario, group = scenario)) +
  geom_line(size = 1) +
  labs(title = NULL, 
       x = "Year", y = "kgco2e/kWh", color = "Scenario") +
  scale_color_manual(
    values = p4_colors,
    breaks = p4_scenario_order) +
  theme_minimal() +
  theme(text = element_text(size = 8, color = "black"),
        plot.title = element_text(face = "bold", size = 12, color = "black"),
        axis.text = element_text(size = 8, color = "black"),
        axis.title = element_text(size = 8, color = "black"),
        legend.text = element_text(size = 8, color = "black"),
        axis.ticks = element_line(color = "black", size = .5),
        axis.line = element_line(color = "black", size = .5),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        aspect.ratio = .5,
        legend.position = "right")

ggsave("figures/fig4.png", plot = p4, width = 6.5, height = 2.5, dpi = 300)

```

Figure 5: 6 panel CBA indicators
```{r}
# Calculates three seperate tables of NPV for private, economic and social costs, using the range of discount rates provided
dr_lb <- 1
dr_ub <- 8

cba_discountrange(dr_min = dr_lb, dr_max = dr_ub, scenario_lifetime = scenario_years)

# Calculates statistics for the Monte Carlo Simulation (lower bound, median, and upper bound)

mcs_cba_stats(lower_bound_out = .05, lower_bound_in = .25, upper_bound_in = .75, upper_bound_out = .95)

# need to convert discount rate column to numeric
npv_stats$dr <- as.numeric(npv_stats$dr) 
sroi_stats$dr <- as.numeric(sroi_stats$dr) 
lcca_stats$dr <- as.numeric(lcca_stats$dr) 

# Produce each panel

cba_econ_plot("npv_econ_plot", npv_stats, c = "orangered4", ylab = "NPV (USD)\n ", xlab = "r", panel = "a")
cba_social_plot("npv_social_plot", npv_stats, c = "orangered4", ylab = NULL, xlab = "r", panel = "b")
cba_econ_plot("sroi_econ_plot", sroi_stats, c = "deepskyblue4", ylab = "SROI (benefits:costs)\n ", xlab = "r", panel = "c")
cba_social_plot("sroi_social_plot", sroi_stats, c = "deepskyblue4", ylab = NULL, xlab = "r", panel = "d")
cba_econ_plot("lcca_econ_plot", lcca_stats, c = "seagreen4", ylab = "LCCA (cost/tco2e)\n ", xlab = "r\n \nEconomic Boundary", panel = "e")
cba_social_plot("lcca_social_plot", lcca_stats, c = "seagreen4", ylab = NULL, xlab = "r\n \nSocial Boundary", panel = "f")

p5 <- npv_econ_plot + npv_social_plot + sroi_econ_plot + sroi_social_plot + lcca_econ_plot + lcca_social_plot + 
  plot_layout(ncol = 2, nrow = 3)

ggsave("figures/fig5.png", plot = p5, width = 6.5, height = 6.5, dpi = 300)

```