---
title: Monte Carlo Simulation for GHG Measure CBA
author: Brandon
project: Dissertation chapter 3
date: Jan 31, 2025
---

TO-DO:
1. Add description here. Boundary acts like a single functional unit. Each parameter is a single functional unit with multiple GHG, non-GHG, and cost factors.

2. Consider how to deal with scope 3 EFs. Electricity for example, has upstream factors. Combine into one electricity EF? Or keep separate for in/out of jurisdiction reporting? 

3. add costs to excel export. Think about how to export discounted costs

4. Turn each MCS loop into a function. A different function for the different PDF types

Package Install & Load
```{r setup}
if (!require("dplyr")) install.packages("dplyr")
library(dplyr)
if (!require("readxl")) install.packages("readxl")
library(readxl)
if (!require("writexl")) install.packages("writexl")
library(writexl)
if (!require("ggplot2")) install.packages("ggplot2")
library(ggplot2)
if (!require("scales")) install.packages("scales")
library(scales)
if (!require("data.table")) install.packages("data.table")
library(data.table)
```

Set up global variables
```{r}
# Define the following global variables to be called by parameter functions
run_count <- 1000
scenario_start <- 2025
scenario_end  <- 2050
year_count <- scenario_end - scenario_start + 1
output_headers <- scenario_start:scenario_end
```

Initiate helper functions
```{r}
source("parameter_functions.r")
```

Create CO2e Emission Factor Library
```{r}
EFL.co2e("ar6")
write_xlsx(EFL, path = "EFLco2e.xlsx")
```

Emission Factor Simulations
```{r}
scenario.MCS("electricity_ef")
```

Parameter 1 Boundary Simulations
```{r}
# Reference boundary represents zero homes funded each year
static.value("ref_boundary", 0)

# Intervention boundary is modeled as a random value between 18 and 21 for out through 2035 (11 years), then zero home per year afterwards.
random.value("intv_boundary", minvalue = 18, maxvalue = 21, lifetime = 11, value2 = 0)

# Transform boundary data from annual totals to cumulative totals

ref_boundary_cumsum <- as.data.frame(t(apply(ref_boundary, 1, cumsum))) # cumulative reference boundary.
intv_boundary_cumsum <- as.data.frame(t(apply(intv_boundary, 1, cumsum))) # cumulative intervention boundary
net_boundary_cumsum <- intv_boundary_cumsum-ref_boundary_cumsum # net cumulative boundary

```

Parameter 1 Factor Simulations
```{r}
# Reference factor 1 represents the annual electricity consumption for a residential home, assumed to be 9823 kwh
static.value("ref_factor1", 9823)

# Reference factor 2 represents the annual natural gas consumption for a residential home, assumed to be 118 therms
static.value("ref_factor2", 118)

# Intervention factor 1 represents energy savings potential of the measure, provided as a % reduction in energy use. 
# The energy savings is applied to both electricity and natural gas
# The energy savings is modeled as a normal distribution: mean = 30% sd = 3%
# The selected value for each simulation is applied consistently over the scenario lifetime.
normal.PDF("intv_factor1", .3, .03)

# Calculate cumulative activity data for factor 1 (electricity, kWh)
calculate.AD("ref_AD1", "intv_AD1", net_boundary_cumsum, ref_factor1, intv_factor1, "kWh")

# Calculate cumulative activity data for factor 2 (natural gas, therms)
calculate.AD("ref_AD2", "intv_AD2", net_boundary_cumsum, ref_factor2, intv_factor1, "therms")
```

Parameter 1 Cost Simulations
```{r}
# Reference capex1 is $0 per home
static.value("ref_capex1_costfactor", 0)

# Reference opex1 is $0.16 per kWh electricity
static.value("ref_opex1_costfactor", -.16)

# Reference opex2 is $2.40 per therm natural gas
static.value("ref_opex2_costfactor", -2.4)

# Intervention capex1 is the program cost per home, modeled as a normal distribution.
# mean = $5117 sd = $52
# The selected value for each simulation is applied consistently over the scenario lifetime.
normal.PDF("intv_capex1_costfactor", -5117, 52)

# calculate reference costs
ref_capex1 <- ref_boundary*ref_capex1_costfactor
ref_opex1 <- ref_AD1[, -1]*ref_opex1_costfactor
ref_opex2 <- ref_AD2[, -1]*ref_opex2_costfactor

# calculate intervention costs
intv_capex1 <- intv_boundary*intv_capex1_costfactor
intv_opex1 <- intv_AD1[, -1]*ref_opex1_costfactor # the cost for electricity is the same in the ref & intv scenarios
intv_opex2 <- intv_AD2[, -1]*ref_opex2_costfactor # the cost for natural gas is the same in the ref & intv scenarios

# Consolidate total costs
ref_totalcost <- (ref_capex1 + ref_opex1 + ref_opex2)
intv_totalcost <- (intv_capex1 + intv_opex1 + intv_opex2)

```

Parameter 1 GHG Calculations
```{r}
# calculate GHG emissions for AD1 (electricity)
ghg_conversion2("ref_ghg1", "intv_ghg1", electricity_ef, ref_AD1, intv_AD1)

# calculate GHG emissions for AD2 (natural gas)
ghg_conversion1("ref_ghg2",
                "intv_ghg2",
                ref_AD2,
                intv_AD2,
                fltr.ef_activeyear = 2024, 
                fltr.service_type = "natural gas", 
                fltr.emission_category = "stationary", 
                fltr.emission_scope = "scope 1",
                fltr.service_subcategory1 = "", 
                fltr.service_subcategory2 = "", 
                fltr.country = "global", 
                fltr.subregion = "")

# calculate total reference GHG emissions
ref_ghgtotal <- (ref_ghg1[, -1] + ref_ghg2[, -1])/1000
ref_ghgtotal$unit <- "mtco2e"
ref_ghgtotal <- ref_ghgtotal[, c("unit", setdiff(names(ref_ghgtotal), "unit"))]

# calculate total intervention GHG emissions
intv_ghgtotal <- (intv_ghg1[, -1] + intv_ghg2[, -1])/1000
intv_ghgtotal$unit <- "mtco2e"
intv_ghgtotal <- intv_ghgtotal[, c("unit", setdiff(names(intv_ghgtotal), "unit"))]

# Calculate GHG mitigation
ghg_mitigation <- ref_ghgtotal[, -1] - intv_ghgtotal[, -1]
ghg_mitigation$unit <- "mtco2e"
ghg_mitigation <- ghg_mitigation[, c("unit", setdiff(names(ghg_mitigation), "unit"))]
```


Parameter 1 Co-benefits 
```{r}
# Social cost of carbon (SCC) is modeled as a normal distribution. It's modeled as an avoided cost, so a positive value is used. 
# mean = $44 sd = $37
# The selected value for each simulation is applied consistently over the scenario lifetime.
normal.PDF("coben1", 44, 37)

# The function above leads to negative values for the social cost of carbon. In practice, this does not make sense. So negative values are converted to 0 below
coben1[coben1 < 0] <- 0

# Calculate social value from Parameter 1 co-benefits 

coben1_value <- ghg_mitigation[, -1] * coben1
coben1_value$unit <- "social value of avoided carbon (USD)"
coben1_value <- coben1_value[, c("unit", setdiff(names(coben1_value), "unit"))]
total_coben_value <- coben1_value[, -1]

```

Parameter 1 cashflow outputs
```{r}
cashflow_private <- intv_capex1 - ref_capex1 # assuming capital expenses are the only private cost/benefit
cashflow_econ <- intv_totalcost - ref_totalcost
cashflow_social <- cashflow_econ + total_coben_value
```

Excel Export
```{r}
p1_outputlist <- (list("ref_boundary" = ref_boundary,
                    "intv_boundary" = intv_boundary, 
                    "ref_factor1" = ref_factor1,
                    "ref_factor2" = ref_factor2,
                    "intv_factor" = intv_factor1,
                    "ref_capex1" = ref_capex1,
                    "intv_capex1" = intv_capex1,
                    "ref_opex1" = ref_opex1,
                    "intv_opex1" = intv_opex1,
                    "ref_opex2" = ref_opex2,
                    "intv_opex2" = intv_opex2,
                    "ref_data1" = ref_AD1,
                    "intv_data1" = intv_AD1,
                    "ref_data2" = ref_AD2,
                    "intv_data2" = intv_AD2,
                    "emission_factors_1" = electricity_ef,
                    "ref_ghg1" = ref_ghg1,
                    "intv_ghg1" = intv_ghg1,
                    "ref_ghg2" = ref_ghg2,
                    "intv_ghg2" = intv_ghg2,
                    "ref_ghgtotal" = ref_ghgtotal,
                    "intv_ghgtotal" = intv_ghgtotal,
                    "net_ghg" = ghg_mitigation,
                    "scc" = coben1,
                    "social_value" = coben1_value,
                    "cashflow_private" = cashflow_private,
                    "cashflow_econ" = cashflow_econ,
                    "cashflow_social" = cashflow_social))

#write excel and save p1 outputs (NOTE: Change names for each parameter)

write_xlsx(p1_outputlist, path = "p1_output.xlsx")

```

CBA Indicator calcs with discount rate range and Monte Carlo Minulation statistics
```{r}
# Calculates three seperate tables of NPV for private, economic and social costs, using the range of discount rates provided

cba_discountrange(dr_min = 0, dr_max = 7, scenario_lifetime = year_count)

# Calculates statistics for the Monte Carlo Simulation (lower bound, median, and upper bound)

mcs_stats(lower_bound = .25, upper_bound = .75)

```

NPV Matrix
```{r}
# Econ NPV matrix
npv_econ_medians <- data.frame(matrix(NA, nrow = 26, ncol = 71)) 
drcolnames <- paste0("dr",((0:70)/10))
colnames(npv_econ_medians) <- drcolnames # column headers for each discount rate: seq(0, 7, by = 0.1)
rownames(npv_econ_medians) <- years

# Social NPV matrix
npv_social_medians <- data.frame(matrix(NA, nrow = 26, ncol = 71)) 
colnames(npv_social_medians) <- drcolnames # column headers for each discount rate: seq(0, 7, by = 0.1)
rownames(npv_social_medians) <- years

# Create a cumulative sum of co-benefit value 
coben_cumsum <- as.data.frame(t(apply(total_coben_value, 1, cumsum)))
median_coben_value <- data.frame(t(apply(coben_cumsum, 2, median))) # The median cumulative NPV value for each year is taken
median_coben_value_vector <- as.vector(unlist(median_coben_value))

for (i in 0:70) { # the loop repeats for each discount rate seq(0, 7, by = 0.1)
  drvalues <- numeric(26) # an empty vector 'drvalues' is created to house the loop outputs
  drvalues[1] <- 1        
    for (n in 2:26) {
         drvalues[n] <- (1+(i/1000))^(n-1) # drvalues is populated with the discount rate conversion factor for each year
    }
  drvalues_df <- data.frame(t(drvalues)) # make a dataframe out of the discount rate loop output
  discount_df <- drvalues_df[rep(1, run_count),] # the discount rate vector is replicated into rows equal to the MCS run count
  annual_npv_econ <- cashflow_econ/discount_df # the total annual economic cash flow is converted into NPV
  annual_npv_econ_cumsum <- as.data.frame(t(apply(annual_npv_econ, 1, cumsum))) # npv is summed cumulatively for each year
  median_npv_econ <- data.frame(t(apply(annual_npv_econ_cumsum, 2, median))) # median cumulative NPV value is taken for each year
  median_npv_vector <- as.vector(unlist(median_npv_econ))
  median_npv_social <- median_npv_vector + median_coben_value_vector# calculate the social npv
  npv_econ_medians[,i+1] <- median_npv_vector #bind the economic NPV values for the given discount rate
  npv_social_medians[,i+1] <- median_npv_social #bind the NPV social values for the given discount rate
}

```

Plot: Annual GHG Mitigation
```{r}

median_mitigation <- as.data.frame(t(apply(p1_ghg_mitigation, 2, median)))
cumulative_mitigation <- as.data.frame(t(apply(median_mitigation, 1, cumsum)))
mitigation_plot_data <- data.frame(
  Year1 = 2025:2050,
  Annual1 = unlist(median_mitigation[1,]),
  Cumulative1 = unlist(cumulative_mitigation[1,])
)

# Create the chart with a legend for cumulative and annual data
ggplot(mitigation_plot_data, aes(x = Year1)) +
  # Bar chart for cumulative data
  geom_bar(aes(y = Cumulative1, fill = "Cumulative Mitigation"), 
           stat = "identity", color = "black", width = 0.4, alpha = 0.7) +
  # Line chart for annual data
  geom_line(aes(y = Annual1 / max(Annual1) * max(Cumulative1), linetype = "Annual Mitigation"), 
            color = "black", linewidth = 1) +
  # Scale the primary y-axis and format with commas
  scale_y_continuous(
    name = "Cumulative GHG Reductions (mtco2e)",
    labels = label_comma(),  # Apply comma formatting to primary y-axis
    #breaks = seq(0, max(mitigation_plot_data$Cumulative1), by = 100),  # Adjust the 'by' parameter to control spacing
    sec.axis = sec_axis(
      ~ . * max(mitigation_plot_data$Annual1) / max(mitigation_plot_data$Cumulative1),
      name = "Annual GHG Reductions (mtco2e)",
      labels = label_comma()  # Apply comma formatting to secondary y-axis
    )
  ) +
  # Customize labels and theme
  labs(
    title = "Estimate of GHG Mitigation",
    x = "Year"
  ) +
  scale_fill_manual(values = c("Cumulative Mitigation" = "green4")) +  # Color for cumulative bars
  scale_linetype_manual(values = c("Annual Mitigation" = "dashed")) +  # Line style for annual data
  theme_minimal() +
  theme(
    legend.position = "bottom",  # Position the legend at the top
    legend.title = element_blank()  # Remove the legend title (Data Type)
  ) +
  guides(
    fill = guide_legend(title = NULL),  # Remove the legend title for fill (bars)
    linetype = guide_legend(title = NULL)  # Remove the legend title for linetype (line)
    )
```

Plot: NPV Histograms
```{r}
# Histogram of Economic NPV with 3% discount rate
npv_econ <- as.data.frame(npv_econ_dr[,30])
npv_econ_median <- npv_econ_stats[31,2]
npv_econ_lb <- npv_econ_stats[31,3]
npv_econ_ub <- npv_econ_stats[31,4]

ggplot(npv_econ, aes(x = npv_econ[,1])) +
  geom_histogram(binwidth = 20000, fill = "darkgrey", color = "black", alpha = 0.7) +
  geom_vline(aes(xintercept = npv_econ_median), color = "red", linetype = "dashed", size = 1) +  # median line
  geom_vline(aes(xintercept = npv_econ_lb), color = "green", linetype = "dotted", size = 1) +  # lower bound
  geom_vline(aes(xintercept = npv_econ_ub), color = "green", linetype = "dotted", size = 1) +  # upper bound
  labs(title = "Histogram of Economic NPV, with median and 25-75% percentiles", x = "npv", y = "Frequency") +
  scale_x_continuous(breaks = seq(-1000000, 1000000, by = 100000), labels = scales::comma) # need to adjust scales manually

# Histogram of Social NPV with 3% discount rate
npv_social <- as.data.frame(npv_social_dr[,30])
npv_social_median <- npv_social_stats[31,2]
npv_social_lb <- npv_social_stats[31,3]
npv_social_ub <- npv_social_stats[31,4]

ggplot(npv_social, aes(x = npv_social[,1])) +
  geom_histogram(binwidth = 20000, fill = "darkgrey", color = "black", alpha = 0.7) +
  geom_vline(aes(xintercept = npv_social_median), color = "red", linetype = "dashed", size = 1) +  # median line
  geom_vline(aes(xintercept = npv_social_lb), color = "green", linetype = "dotted", size = 1) +  # lower bound
  geom_vline(aes(xintercept = npv_social_ub), color = "green", linetype = "dotted", size = 1) +  # upper bound
  labs(title = "Histogram of Social NPV, with median and 25-75% percentiles", x = "npv", y = "Frequency") +
  scale_x_continuous(breaks = seq(-1000000, 1000000, by = 100000), labels = scales::comma) # need to adjust scales manually

# Overlayed NPV Histograms
set.seed(123)
npv1 <- npv_econ[,1]
npv2 <- npv_social[,1]

# Combine the two datasets into a single data frame
npvplotdf <- data.frame(
  value = c(npv1, npv2),
  Dataset = rep(c("Economic LCCA", "Social LCCA"), each = 1000)
)

# Plot two histograms on top of each other
ggplot(npvplotdf, aes(x = value, fill = Dataset)) +
  geom_histogram(alpha = 0.5, position = "identity", bins = 30) +
  geom_vline(aes(xintercept = npv_econ_median), color = "darkblue", linetype = "dashed", size = 1) +  # median line
  geom_vline(aes(xintercept = npv_social_median), color = "darkgreen", linetype = "dashed", size = 1) +  # lower bound
  scale_fill_manual(values = c("darkblue", "darkgreen")) +  # Custom colors for the two datasets
  labs(title = "Overlayed NPV Histograms with Median on Dashed Line", x = "NPV", y = "Frequency") +
  scale_x_continuous(breaks = seq(-1000000, 1000000, by = 200000), labels = scales::comma) # need to adjust scales manually

```

Plot: NPV Based on discount rate
```{r, eval = FALSE}
# Economic NPV
# Convert years to numeric for plotting
npv_econ_medians$year <- years

# Convert discount rate to numeric for plotting
npv_econ_medians$year <- as.numeric(npv_econ_medians$year)

# plot
ggplot(npv_econ_medians, aes(x = year, y = dr3)) +
  geom_line(color = "darkblue", size = 1) +                      # Line for the mean
  geom_ribbon(aes(ymin = dr0, ymax = dr7), fill = "darkblue", alpha = 0.2) +  # Shaded confidence interval
  labs(title = "NPV at a 3% Discount rate, with 0% - 7% discount rate shaded",
       x = "Year",
       y = "USD") +
  scale_y_continuous(breaks = seq(-1000000, 1000000, by = 200000),labels = comma) + # Format y-axis with commas
  theme_minimal()

# Scoial NPV
# Convert years to numeric for plotting
npv_social_medians$year <- years

# Convert discount rate to numeric for plotting
npv_social_medians$year <- as.numeric(npv_social_medians$year)

# plot
ggplot(npv_social_medians, aes(x = year, y = dr3)) +
  geom_line(color = "darkgreen", size = 1) +                      # Line for the mean
  geom_ribbon(aes(ymin = dr0, ymax = dr7), fill = "darkgreen", alpha = 0.2) +  # Shaded confidence interval
  labs(title = "NPV at a 3% Discount rate, with 0% - 7% discount rate shaded",
       x = "Year",
       y = "USD") +
  scale_y_continuous(breaks = seq(-1000000, 1000000, by = 200000),labels = comma) + # Format y-axis with commas
  theme_minimal()

# Overlayed Line Plot
# New datasets for Overlay
econnpv_overlay <- npv_econ_medians
socialnpv_overlay <- npv_social_medians

# Add group name
econnpv_overlay$group <- "Economic NPV"
socialnpv_overlay$group <- "Social NPV"

# Combine datasets
npv_overlayline_data <- rbind(econnpv_overlay, socialnpv_overlay)

# Create the plot
ggplot(npv_overlayline_data, aes(x = year, y = dr3, group = group)) +
  geom_ribbon(aes(ymin = dr2, ymax = dr4, fill = group), alpha = 0.2) +
  geom_line(aes(color = group), size = 1) +
  scale_fill_manual(values = c("Economic NPV" = "darkblue", "Social NPV" = "darkgreen")) +
  scale_color_manual(values = c("Economic NPV" = "darkblue", "Social NPV" = "darkgreen")) +
  theme_minimal() +
  scale_y_continuous(breaks = seq(-1000000, 1000000, by = 200000),labels = comma) + # Format y-axis with commas
  labs(title = "NPV at a 3% Discount rate, with 2% - 4% discount rate shaded",
       x = "Year",
       y = "USD",
       fill = "npv_overlayline_data",
       color = "npv_overlayline_data")

```

Plot: LCCA Histograms
```{r}
# Histogram of Economic NPV with 3% discount rate
lcca_econ <- as.data.frame(lcca_econ_dr[,30])
lcca_econ_median <- lcca_econ_stats[31,2]
lcca_econ_lb <- lcca_econ_stats[31,3]
lcca_econ_ub <- lcca_econ_stats[31,4]

ggplot(lcca_econ, aes(x = lcca_econ[,1])) +
  geom_histogram(binwidth = 20, fill = "blue", color = "black", alpha = 0.7) +
  geom_vline(aes(xintercept = lcca_econ_median), color = "red", linetype = "dashed", size = 1) +  # median line
  geom_vline(aes(xintercept = lcca_econ_lb), color = "green", linetype = "dotted", size = 1) +  # lower bound
  geom_vline(aes(xintercept = lcca_econ_ub), color = "green", linetype = "dotted", size = 1) +  # upper bound
  labs(title = "Histogram of Economic lcca, with median and 25-75% percentiles", x = "lcca", y = "Frequency") +
  scale_x_continuous(breaks = seq(-500, 500, by = 50), labels = scales::comma) # need to adjust scales manually

# Histogram of Social lcca with 3% discount rate
lcca_social <- as.data.frame(lcca_social_dr[,30])
lcca_social_median <- lcca_social_stats[31,2]
lcca_social_lb <- lcca_social_stats[31,3]
lcca_social_ub <- lcca_social_stats[31,4]

ggplot(lcca_social, aes(x = lcca_social[,1])) +
  geom_histogram(binwidth = 20, fill = "blue", color = "black", alpha = 0.7) +
  geom_vline(aes(xintercept = lcca_social_median), color = "red", linetype = "dashed", size = 1) +  # median line
  geom_vline(aes(xintercept = lcca_social_lb), color = "green", linetype = "dotted", size = 1) +  # lower bound
  geom_vline(aes(xintercept = lcca_social_ub), color = "green", linetype = "dotted", size = 1) +  # upper bound
  labs(title = "Histogram of Social lcca, with median and 25-75% percentiles", x = "lcca", y = "Frequency") +
  scale_x_continuous(breaks = seq(-500, 500, by = 50), labels = scales::comma) # need to adjust scales manually

# Overlayed lcca Histograms
set.seed(123)
lcca1 <- lcca_econ[,1]
lcca2 <- lcca_social[,1]

# Combine the two datasets into a single data frame
lccaplotdf <- data.frame(
  value = c(lcca1, lcca2),
  Dataset = rep(c("Economic LCCA", "Social LCCA"), each = 1000)
)

# Plot two histograms on top of each other
ggplot(lccaplotdf, aes(x = value, fill = Dataset)) +
  geom_histogram(alpha = 0.5, position = "identity", bins = 30) +
  geom_vline(aes(xintercept = lcca_econ_median), color = "darkblue", linetype = "dashed", size = 1) +  # median line
  geom_vline(aes(xintercept = lcca_social_median), color = "darkgreen", linetype = "dashed", size = 1) +  # lower bound
  scale_fill_manual(values = c("darkblue", "darkgreen")) +  # Custom colors for the two datasets
  labs(title = "Overlayed LCCA Histograms with Median on Dashed Line", x = "Value", y = "Frequency") +
  scale_x_continuous(breaks = seq(-500, 500, by = 100), labels = scales::comma) # need to adjust scales manually

```

Plot: LCCA Based on discount rate
```{r, eval = FALSE}
# Economic LCCA
# Convert discount rate to numeric for plotting
lcca_econ_stats$DR <- as.numeric(lcca_econ_stats$DR)

# Filter the confidence results to get rid of any years with zero lcca
plot2_data <- subset(lcca_econ_stats, Median != 0)

# plot
ggplot(plot2_data, aes(x = DR, y = Median)) +
  geom_line(color = "darkblue", size = 1) +                      # Line for the mean
  geom_ribbon(aes(ymin = Lower, ymax = Upper), fill = "darkblue", alpha = 0.2) +  # Shaded confidence interval
  labs(title = "Median, with 25-75% percentile range shaded",
       x = "Discount Rate",
       y = "LCCA") +
  theme_minimal()

# Scoial LCCA
# Convert discount rate to numeric for plotting
lcca_social_stats$DR <- as.numeric(lcca_social_stats$DR)

# Filter the confidence results to get rid of any years with zero lcca
plot3_data <- subset(lcca_econ_stats, Median != 0)

# plot
ggplot(plot3_data, aes(x = DR, y = Median)) +
  geom_line(color = "darkgreen", size = 1) +                      # Line for the mean
  geom_ribbon(aes(ymin = Lower, ymax = Upper), fill = "darkgreen", alpha = 0.2) +  # Shaded confidence interval
  labs(title = "Median, with 25-75% percentile range shaded",
       x = "Discount Rate",
       y = "LCCA") +
  theme_minimal()

# Overlayed Line Plot
# Convert discount rate to numeric for plotting
lcca_econ_stats$DR <- as.numeric(lcca_econ_stats$DR)
lcca_social_stats$DR <- as.numeric(lcca_social_stats$DR)

# Filter the confidence results to get rid of any years with zero lcca
econlcca_linedata <- subset(lcca_econ_stats, Median != 0)
sociallcca_line_data <- subset(lcca_social_stats, Median != 0)

# Add group name
econlcca_linedata$group <- "Economic LCCA"
sociallcca_line_data$group <- "Social LCCA"

# Combine datasets
overlayline_data <- rbind(econlcca_linedata, sociallcca_line_data)

# Create the plot
ggplot(overlayline_data, aes(x = DR, y = Median, group = group)) +
  geom_ribbon(aes(ymin = Lower, ymax = Upper, fill = group), alpha = 0.2) +
  geom_line(aes(color = group), size = 1) +
  scale_fill_manual(values = c("Economic LCCA" = "darkblue", "Social LCCA" = "darkgreen")) +
  scale_color_manual(values = c("Economic LCCA" = "darkblue", "Social LCCA" = "darkgreen")) +
  theme_minimal() +
  labs(title = "Median LCCA, with 25-75% percentile range shaded",
       x = "Discount Rate",
       y = "LCCA",
       fill = "overlayline_data",
       color = "overlayline_data")

```